Variational Inference for Inverse Reinforcement Learning with Gaussian Processes

### 2019-03-15 mpaper/paper.tex
A bit of progress on the proof

\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\adj}{adj}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator*{\argmax}{arg\,max}

@@

\begin{lemma} \label{lemma:bound1}
  Let $c : \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to (a, b) \subset
  \mathbb{R}$ be an arbitrary bounded function. Then, for $i = 0,
  \dots, d$,
  \[
    \left. \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} \right|_{\lambda_i
      = c(\mathbf{r}, \mathbf{u})}
  \]
  has upper and lower bounds of the form $q(\mathbf{r} \mid \mathbf{u})d(\mathbf{u})$, where
  $d(\mathbf{u}) \in \mathbb{R}_2[\mathbf{u}]$.
\end{lemma}
\begin{proof}
  Remember that
  \begin{align*}
    \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} &= \frac{1}{2}q(\mathbf{r} \mid \mathbf{u}) (|\bm\Gamma|^{-1} \tr(\mathbf{R} \adj(\bm\Gamma)) \\
                                                                      &- (\mathbf{r} - \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r} - \mathbf{Su})).
  \end{align*}
  by Lemma \ref{lemma:derivatives}.

  Given any covariance matrix $\mathbf{K}$ and $i = 0, \dots, d$, we can easily
  establish constant upper and lower bounds on $\mathbf{K}|_{\lambda_i =
    c(\mathbf{r}, \mathbf{u})}$ by the boundedness of $c$ and continuity of the
  covariance function. Similar reasoning combined with the expressions for
  derivatives of covariance matrices in Lemma \ref{lemma:derivatives} gives
  constant upper and lower bounds on the elements of $\left.
    \frac{\mathbf{\partial K}}{\partial \lambda_i} \right|_{\lambda_i =
    c(\mathbf{r}, \mathbf{u})}$ as well. We can also use Lemma ?? to show
  constant upper and lower bounds on $\mathbf{K}^{-1}|_{\lambda_i =
    c(\mathbf{r}, \mathbf{u})}$. This is enough to prove constant upper and
  lower bounds on $\mathbf{S}$, $\bm\Gamma$, and $\mathbf{R}$, which means that
  $(\mathbf{r} -
  \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r} -
  \mathbf{Su})$ has upper and lower bounds in $\mathbb{R}_2[\mathbf{u}]$.
  % TODO: write the lemma

>>>>

\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\adj}{adj}
%\DeclareMathOperator{\dett}{det}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator*{\argmax}{arg\,max}

@@

% TODO: add this lemma to the graph
\begin{lemma} \label{lemma:matrix_inverse}
  TODO: constant bounds on the inverse
\end{lemma}

\begin{lemma} \label{lemma:bound1}
  Let $i \in \{ 0, \dots, d \}$ and $\epsilon > 0$ be arbitrary. Furthermore,
  let $c : \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to (\lambda_i -
  \epsilon, \lambda_i + \epsilon) \subset \mathbb{R}$ be a function with a
  codomain arbitrarily close to $\lambda_i$. Then
  \[
    \left. \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i}
    \right|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}
  \]
  has upper and lower bounds of the form $q(\mathbf{r} \mid
  \mathbf{u})d(\mathbf{u})$, where $d(\mathbf{u}) \in \mathbb{R}_2[\mathbf{u}]$.
\end{lemma}
\begin{proof}
  Remember that
  \begin{align*}
    \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} &= \frac{1}{2}q(\mathbf{r} \mid \mathbf{u}) (|\bm\Gamma|^{-1} \tr(\mathbf{R} \adj(\bm\Gamma)) \\
                                                                      &- (\mathbf{r} - \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r} - \mathbf{Su})).
  \end{align*}
  by Lemma \ref{lemma:derivatives}.

  Given any covariance matrix $\mathbf{K}$, we can easily establish constant
  upper and lower bounds on $\mathbf{K}|_{\lambda_i = c(\mathbf{r},
    \mathbf{u})}$ by the boundedness of $c$ and continuity of the covariance
  function. Similar reasoning combined with the expressions for derivatives of
  covariance matrices in Lemma \ref{lemma:derivatives} gives constant upper and
  lower bounds on the elements of
  \[
    \left. \frac{\mathbf{\partial K}}{\partial \lambda_i} \right|_{\lambda_i =
      c(\mathbf{r}, \mathbf{u})}
  \]
  as well. We can also use Lemma \ref{lemma:matrix_inverse} to show constant
  upper and lower bounds on $\mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r},
    \mathbf{u})}$. This is enough to prove constant upper and lower bounds on
  $\mathbf{S}$, $\bm\Gamma$, and $\mathbf{R}$ (all with $\lambda_i$ replaced
  with $c(\mathbf{r}, \mathbf{u})$), which means that $(\mathbf{r} -
  \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r} -
  \mathbf{Su})|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ has upper and lower
  bounds in $\mathbb{R}_2[\mathbf{u}]$.

  Since $\bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ is bounded, so is
  $\det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$. Assuming that
  $\bm\Gamma$ is invertible so that $q(\mathbf{r} \mid \mathbf{u})$ exists, we
  can make $\bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ close enough to
  $\bm\Gamma$ to ensure that
  \[
    \det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \ne 0.
  \]
  Moreover, boundedness of $\bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$
  gives boundedness of $\det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r},
    \mathbf{u})}$.

  Combined with the fact that $ab > 0$ (i.e., both endpoints together are either
  positive or negative), this gives boundedness of $\det(\bm\Gamma)|_{\lambda_i
    = c(\mathbf{r}, \mathbf{u})}^{-1}$.

  TODO: Show that $\bm\Gamma^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ is
  bounded.

  Since we already know that $\bm\Gamma^{-1}|_{\lambda_i = c(\mathbf{r},
    \mathbf{u})}$ and $\det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$
  are bounded, so is $\adj(\bm\Gamma) =
  \det(\bm\Gamma)\bm\Gamma^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$. Thus,
  we have constant bounds on
  $|\bm\Gamma|^{-1}\tr(\mathbf{R}\adj(\bm\Gamma))|_{\lambda_i = c(\mathbf{r},
    \mathbf{u})}$, which means that
  \[
    \left. \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i}
    \right|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}
  \]
  has the required bounds.
